---
title: "G-test"
author: "Emily Yang"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## G–test of goodness-of-fit

Like the Chi-squared test in Module 14 the G test allows you to determine if there is a significant difference in proportions of a categorical variable to the theoretical expectation. Chi-squared test give approximately the same results as G test but you can perform more elaborate statistical analyses with G test results. Downside is that less people are familar with the G test compared to the Chi-squared test. You cannot use both and pick the one that gives you a better result. Pick one and stick with it. 

Both G test and Chi-squared Test data needs to fulfill certain assumptions: 

1. Random Sample <br>
2. Independence between samples <br>
3. Variable must be proportional or categorical <br>
4. Mutually exclusive categories <br>
5. Variables must have 2 or more options <br>


You should use the G test over the Chi-squared Test if 

1. There are more than 1000 values in total.
2. There are outliers in the data

Null Hypothesis: The observed data has the same porportions as the expected theoretical population. 
Alternate Hypothesis: The observed data doesn't have the same porportions as the expected theoretical population. 

## Running the Test

``` {r Generating Data, include=FALSE}

set.seed(1)
Flower.color <- sample(x = c("purple", "white", "orange", "blue"), prob= c(.5, .35, 0.05, 0.1), size = 1200, replace = TRUE)

Flower.size <- sample(x = c("Large", "Small"), prob= c(0.5, 0.5), size = 1200, replace = TRUE)
table(Flower.size) # Count how many of each size

Flower.species <- sample(x = c("Siberian", "Japanese", "Dutch", "Bearded"), prob= c(.25, .25, 0.25, 0.25), size = 1200, replace = TRUE)
table(Flower.species) # Count how many of each species

Data <- cbind(Flower.color, Flower.size, Flower.species)
head(Data)
```
Formula <br>

# G=2 * Σ[O * ln(O/E)]

O = Observed
E = Expected

I was out in the field and collected data about 1200 Irises. I want to know if there is proportionally equal amounts of each flower color. 

```{r Tabulating Data}

table(Flower.color) # Count how many of each color

q <- 2 * (121 * log(121/300)) + (59 * log(59/300)) + (628 * log(628/300)) + (392 * log(392/300))

q

pchisq(q=253.1, df=3, lower.tail = FALSE) # Finding the p-value
```
We got an output of 253.1 and the degree of freedom is options-1 and calculated a p-value of 1.4e-54 which is below the critical p-value = 0.05. We can reject the null hypothesis and conclude that the flower color proportions of the Irises I observed are not equally distributed. 

```{r Coding}
library(DescTools)

#perform the G-test 
GTest(x = c(121, 59, 628, 392), #observed values
      p = c(1/4, 1/4, 1/4, 1/4), #expected proportions
      correct = "none") 

```

Once again the p-value is below 0.05. We can reject the null hypothesis and conclude that the flower color proportions of the Irises I observed are not equally distributed. 


## Challenges

Go through the same process with Flower.size and Flower.species.

### Sources
https://www.statology.org/g-test/ <br>
https://www.statstest.com/g-test/ <br>
https://www.biostathandbook.com/gtestgof.html <br>
https://www.r-bloggers.com/2022/05/calculate-the-p-value-from-chi-square-statistic-in-r/ <br>
https://stats.libretexts.org/Bookshelves/Applied_Statistics/Biological_Statistics_(McDonald)/02%3A_Tests_for_Nominal_Variables/2.04%3A_GTest_of_Goodness-of-Fit <br> <br>
